A Constraint Solving Approach to Parikh Images of Regular Languages

### 2023-10-20 evaluation.tex
discussion of data

In \cref{fig:solve-division}, we show how many of the~\NrBenchmarks{} instances
the respective back-end could solve. A summary of their outcomes by instance
type is also available in \cref{tab:solve-status}. Note that many instances lack
a corum ground truth as they are solved by only one backend. We see that
\Calculus{} generally outperforms \Nuxmv{} on determining unsatisfiability, as
does baseline, while being similar at satisfiable instances. Both \Calculus{} and
\Nuxmv{} outperforms baseline on satisfiable instances. On satisfiable
instances, \Nuxmv and \Calculus{} have similar performance.

@@

\begin{table}
  \begin{center}
  \input{experiments/qf_slia_comparison.tex}
  \end{center}
  \caption{Number of solved benchmarks from the single-query QF\_SLIA track of
  quantifier-free string linear integer arithmetic at SMT-COMP 2023. The numbers
  for each solver are from the publicly available data files, except for
  \Ostrich-\Catra{}, which are executed by us. }\label{tab:solve-status-smt-comp}
\end{table}

@@

 At SMT-COMP, \Ostrich{} competed using different
 (automata-based) back-ends:
\begin{itemize}
\item \textbf{BW-Str:} a backward-propagation-based solver, not utilising
  Parikh automata~\cite{ostrich}.
%\item \textbf{ADT-Str:} a solver based on algebraic data-types.
\item \textbf{CE-Str:} a re-implementation of the \OstrichPlus{}
  algorithm~\cite{ostrich-plus}, using Parikh automata and the
  encoding from~\cite{generate-parikh-image}.
\end{itemize}

For our experiments, we modified the CE-Str solver to apply \Catra{}
instead of the previous baseline method, resulting in a new back-end
\textbf{CA-Str}.

As benchmarks, we again focused on the PyEx benchmarks~\cite{pyex}, of
which we picked 1000~problems uniformly at random. We ran
different combinations of the
back-ends on those 1000~problems with a 120s timeout.
%
%The \texttt{QF\_S} track contains no Parikh automata intersection problems,
%however. For that, we need to look to the linear integer arithmetic
%(\texttt{QF\_SLIA}), where \Ostrich{} did worse in SMT-COMP. The evaluations in
%\cref{sec:scaling} use inputs generated by an older branch of CEFA-enriched
%\Ostrich{} on the PyEx benchmarks, which are part of the set used in the
%competition. In this section, we evaluated a version of \Ostrich{} using
%\Catra{} as its back-end \Fudge{on StarExec, the same cluster SMT-COMP ran on}
%on \numprint{1000}~randomly selected instances from the PyEx benchmarks. 
The results are shown in \cref{tab:solve-status-smt-comp}.
%
% comparison \Fudge{to the other provers of the 2023
%competition on the same benchmarks, as well as \Ostrich{} using the baseline
%back-end described in \cref{sec:implementing-baseline}}.

As can be seen from the table, integrating \Catra{} as a back-end
leads to gains both on satisfiable and unsatisfiable problems. The
increase is more relevant for the unsatisfiable problems, given that
\Ostrich{} is a state-of-the-art solver in particular on such benchmarks.

>>>>

In \cref{fig:solve-division}, we show how many of the~\NrBenchmarks{} instances
the respective back-end could solve. A summary of their outcomes by instance
type is also available in \cref{tab:solve-status}. Note that many instances lack
a corum ground truth as they are solved by only one backend. We see that
\Calculus{} generally outperforms \Nuxmv{} on determining unsatisfiability, as
does baseline, while being similar at satisfiable instances. Both \Calculus{} and
\Nuxmv{} outperforms baseline on satisfiable instances. On satisfiable
instances, \Nuxmv{} and \Calculus{} have similar performance.

@@

\begin{table}
  \begin{center}
  \input{experiments/qf_slia_comparison.tex}
  \end{center}
  \caption{Number of solved benchmarks from the single-query QF\_SLIA
    track of quantifier-free string linear integer arithmetic at
    SMT-COMP 2023. The numbers for each solver are from the publicly
    available data files, except for CA-Str, which is executed by us
    (600s timeout, AMD Opteron 2220
    SE), and \Ostrich+CA, which is a virtual portfolio. }
  \label{tab:solve-status-smt-comp}
\end{table}

@@

 At SMT-COMP, \Ostrich{} competed using a portfolio of three different
 back-ends:
\begin{itemize}
\item \textbf{BW-Str:} a backward-propagation-based solver, not utilising
  Parikh automata~\cite{ostrich}.
\item \textbf{ADT-Str:} a solver based on algebraic data-types.
\item \textbf{CE-Str:} a re-implementation of the \OstrichPlus{}
  algorithm~\cite{ostrich-plus}, using Parikh automata and the
  encoding from~\cite{generate-parikh-image}.
\end{itemize}

For our experiments, we modified the CE-Str solver to apply \Catra{}
instead of the previous baseline method, resulting in a new back-end
\textbf{CA-Str}. We created also a new virtual portfolio~\Ostrich{}+CA
by combining the three solvers in \Ostrich{}~1.3 with \textbf{CA-Str}.

%As benchmarks, we again focused on the PyEx benchmarks~\cite{pyex}, of
%which we picked 1000~problems uniformly at random. We ran
%different combinations of the
%back-ends on those 1000~problems with a 120s timeout.
%
%The \texttt{QF\_S} track contains no Parikh automata intersection problems,
%however. For that, we need to look to the linear integer arithmetic
%(\texttt{QF\_SLIA}), where \Ostrich{} did worse in SMT-COMP. The evaluations in
%\cref{sec:scaling} use inputs generated by an older branch of CEFA-enriched
%\Ostrich{} on the PyEx benchmarks, which are part of the set used in the
%competition. In this section, we evaluated a version of \Ostrich{} using
%\Catra{} as its back-end \Fudge{on StarExec, the same cluster SMT-COMP ran on}
%on \numprint{1000}~randomly selected instances from the PyEx benchmarks. 
%
% comparison \Fudge{to the other provers of the 2023
%competition on the same benchmarks, as well as \Ostrich{} using the baseline
%back-end described in \cref{sec:implementing-baseline}}.

We evaluate the different solvers in \cref{tab:solve-status-smt-comp}
on 21\,938~benchmarks from the single-query QF\_SLIA track at SMT-COMP 2023, as
QF\_SLIA was the most challenging track for \Ostrich{} in the
competition.
%
As can be seen from the table, integrating \Catra{} as a back-end
leads to gains both on satisfiable and unsatisfiable problems. On
satisfiable benchmarks, the combination \Ostrich+CA is still
outperformed by cvc5 and z3alpha. On unsatisfiable benchmarks,
\Ostrich+CA narrowly beats the other solvers, which is promising given
that also \Ostrich{}~1.3, cvc5, and z3alpha show very strong
performance on this class of benchmarks.
